\documentclass[a4paper,11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb, amsmath, amsthm, mathrsfs}
\usepackage[left=1.0in,right=1.0in,top=1.0in,bottom=1.0in]{geometry}
\usepackage[inline,shortlabels]{enumitem}
\usepackage{times}
\usepackage{xcolor}

\newcommand{\R}{\mathbf{R}}

\begin{document}

\begin{center}
  {\Large\bfseries Math 240 Tutorial \\ Questions}
\end{center}
\begin{center}
  {\bfseries May 16}
\end{center}

\noindent{\bfseries Question 1.} Place the following augmented matrices into
an echelon form. Does the corresponding system of linear equations admit any
solutions?
\begin{enumerate}[(a)]
\item
  \[
    \left(
      \begin{array}{cccc|c}
        4 & 8 & 12 & 4 & 7 \\
        2 & 5 & 6 & 6 & 11 \\
        0 & 5 & 1 & 26 & 13 \\
        0 & 5 & 0 & 21 & 17
      \end{array}
    \right).
  \]
\item
  \[
    \left(
      \begin{array}{cccc|c}
        4 & 8 & 12 & 4 & 0 \\
        2 & 5 & 6 & 6 & 0 \\
        0 & 5 & 1 & 25 & 0 \\
        0 & 5 & 0 & 20 & 0       
      \end{array}
    \right).
  \]
\item
  \[
    \left(
      \begin{array}{cccc|c}
        4 & 8 & 12 & 4 & 7 \\
        2 & 5 & 6 & 6 & 11 \\
        0 & 5 & 1 & 25 & 13 \\
        0 & 5 & 0 & 20 & 17       
      \end{array}
    \right).
  \]
\end{enumerate}

\noindent{\bfseries Question 2.} Consider the following system of equations
\begin{align*}
  w + x + y + z &= 6, \\
  w + y + z &= 4, \\
  w + y &=2.
\end{align*}
\begin{enumerate}[(a)]
\item List the leading variables.
\item List the free variables.
\item Write the general solution to the equation (expressed in terms of the free
  variables).
\item Suppose a fourth equation $-2w+y=5$ is added to the system. What is the
  solution of the resulting system?
\item Suppose the fourth equation is $-2w-2y=-3$. What can we say about the
  solutions for the resulting system? \\
\end{enumerate}

\noindent{\bfseries Question 3.} Find the values of $k$ for which the system of
equations
\begin{align*}
  x + ky &= 1, \\
  kx + y &= 1,
\end{align*}
has
\begin{enumerate}[(a)]
\item no solution,
\item a unique solution, and
\item infinitely many solutions.
\item When there is exactly one solution, what are the values of $x$ and $y$. \\
\end{enumerate}


\noindent{\bfseries Question 4.} Consider the following system of linear equations
\begin{align*}
  u+2v-w-2x+3y &= b_1, \\
  x-y+2z &= b_2, \\
  2u+4v-2w-4x+7y-4z &= b_3, \\
  -x+y-2z &= b_4, \\
  3u+6v-3w-6x+7y+8z &= b_5,
\end{align*}
where $b_1,\,b_2,\,b_3,\,b_4,\,b_5 \in \R$.
\begin{enumerate}[(a)]
\item What are the leading and free variables?
\item\label{Q2:conditions} What conditions must the real constants
  $b_1,\,b_2,\,b_3,\,b_4,\,b_5$ satisfy in order that the system be consistent?
\item Do the numbers $b_1=1$, $b_2=-3$, $b_3=2$, $b_4=b_5=3$ satisfy the
  conditions of part \ref{Q2:conditions}? If so, find the general solution in
  terms of the free variables.
\end{enumerate}

\end{document}